Breanna_Code

### 2022-10-05 Draft_Report.Rmd
I've updated our initial plots and added in some data for the axinopsida that allows me to use the binomial glm for the proportion. My model selection and interpretation could use some guidance, most likely. There are a variety of models with a variety of assumptions and it is hard to keep track of them all as these are mostly new to me. I appreciate all your help.

```{r load_packages}
library(dplyr)
library(lme4)
library(tidyverse)
library(lattice)
```

@@

@@

@@

#assess model
library(lattice)

# diagnostic plots
plot(modelrichness2)
  # note that heteroscedasticity looks ok for model residuals
  # note that residuals look mostly normally distributed, 
  # with some heavy tails but overall looks good
  # smooth curve is mostly straight, looks like model fit is ok, 
  # there is a bit of a w shaped pattern, however
```

@@

@@

@@

# Compare both models AIC
anova(modeldiversity1, modeldiversity2)
# Conclude that random effects are significant in the model.
# However, the simpler model has lower AIC. 
# Removing random effects would simplify interpretation

@@

# Diagnostic plots
plot(modeldiversity1, type=c("p","smooth"))
  # type p is the points of the scatterplot, 
  # smooth is the locally estimated scatterplot smoothing (loess)
  # note that heteroscedasticity looks ok for model residuals
  # smooth curve is mostly straight, looks like model fit is ok

# QQ normal plot
qqnorm(resid(modeldiversity1), main="normal qq-plot, residuals")
qqline(resid(modeldiversity1))

  # note that residuals look mostly normally distributed in the qqplot, 
  # with some heavy tails on the lower left.

@@

@@

# Diagnostic plots
plot(modelevenness1, type=c("p","smooth"))
  # type p is the points of the scatterplot, 
  # smooth is the locally estimated scatterplot smoothing (loess)
  # note that heteroscedasticity looks ok for model residuals
  # smooth curve is mostly straight, looks like model fit is ok

# QQ normal plot
qqnorm(resid(modelevenness1), main="normal qq-plot, residuals")
qqline(resid(modelevenness1))
  # note that residuals look mostly normally distributed in the qqplot,

@@

>>>>

```{r load_packages}
library(dplyr)
library(lme4)
library(tidyverse)
library(lattice)

@@

### Abundance

Model used:

Generalized linear mixed model fit by maximum likelihood (Laplace Approximation) ['glmerMod']

Family: poisson ( log )

Formula: Abundance \~ Direction + Distance + (1 \| Station)

Data: metrics

Distance and direction are significant. Abundance increases with distance, abundance is higher on the south side.

@@

### Richness

Model used:

Call:

glm(formula = Richness \~ Direction + Distance + Direction:Distance,

family = poisson, data = metrics)

Where all terms are significant, richness increases with increasing distance on the north side, but only increases slightly with distance on the south side. (The difference in response for each direction is characterized by the significant interaction term.)

@@

#assess model
# diagnostic plots
plot(modelrichness2)
  # note that heteroscedasticity looks ok for model residuals
  # note that residuals look mostly normally distributed, 
  # with some heavy tails but overall looks good
  # smooth curve is mostly straight, looks like model fit is ok, 
  # there is a bit of a w shaped pattern, however
```

### Biomass

Model used:

Call:

lm(formula = log_Biomass \~ Direction + Distance, data = metrics)

The biomass does not change with distance, but sites to the north have decreased biomass compared to the south.

@@

### Swartz Dominance Index

Model used:

Call:

glm(formula = Dominance \~ Direction + Distance + Direction:Distance,

family = poisson, data = metrics)

Where all terms are significant, dominance increases with increasing distance on the north side, and decreases with distance on the south side. (The difference in response for each direction is characterized by the significant interaction term.)

@@

### Simpson's Diversity

Model used:

Call:

lm(formula = Diversity \~ Direction + Distance + Direction:Distance,

data = metrics)

Where distance and the interaction with distance and direction terms are significant, diversity increases with increasing distance on the north side, but increases less dramatically with distance on the south side. (The difference in response for each direction is characterized by the significant interaction term.)

@@

# Compare both models AIC
anova(modeldiversity1, modeldiversity2)
# Random effects are significant in the model.
# However, the simpler model has lower AIC. 
# Removing random effects would simplify interpretation

@@

# Diagnostic plots
plot(modeldiversity2)
  # type p is the points of the scatterplot, 
  # smooth is the locally estimated scatterplot smoothing (loess)
  # note that heteroscedasticity looks ok for model residuals
  # smooth curve is mostly straight, looks like model fit is ok
  # note that residuals look mostly normally distributed in the qqplot, 
  # with some heavy tails on the lower left.

@@

### Simpson's Evenness

Model used:

Linear model

Call:

lm(formula = Evenness \~ Direction + Distance + Direction:Distance,

data = metrics)

Where all terms are significant, evenness decreases with increasing distance on the north side, but increases slightly with distance on the south side. (The difference in response for each direction is characterized by the significant interaction term.)

@@

# Diagnostic plots
plot(modelevenness2)
  # type p is the points of the scatterplot, 
  # smooth is the locally estimated scatterplot smoothing (loess)
  # note that heteroscedasticity looks ok for model residuals
  # smooth curve is mostly straight, looks like model fit is ok
  # note that residuals look mostly normally distributed in the qqplot,

@@

### Axinopsida Proportion of Adults with Low Shell Staining

Model used:

Generalized linear mixed model fit by maximum likelihood (Laplace Approximation) ['glmerMod']

Family: binomial ( logit )

Formula: cbind(Shell_Successes, Shell_Trials - Shell_Successes) \~ Direction + Distance + (1 \| Station)

Where all terms are significant, shell staining decreases with increasing distance, decreases from north to south.
